Back-integration likelihood

### 2014-01-29 text/notes/likelihood/likelihood.tex
some minor tweaks

Take ${\bf D}$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take ${\bf W}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame. We have determined that this star was once part of a progenitor system (e.g., satellite galaxy) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy. The position of the satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, where the subscript $s$ refers to the satellite. The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.

We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the tidal radius, $R_{\rm tide}$, and escape velocity, $V_{\rm esc}$ of the satellite, respectively, at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$. In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref. response?].

With these assumptions we can write the joint posterior probability for the observed position of a star, properties of the disrupting satellite, and parent potential given the data as:
\begin{align}
	%p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
	%	\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi)
	p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
		\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi)
\end{align}
where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters. We assume that the observational errors are Gaussian:
\begin{align}
	p(\D^{(i)} \given \W^{(i)}) &= \mathcal{N}(\W^{(i)} \given \D^{(i)}, \bSigma^{(i)})\\
	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bSigma_s)
\end{align}
where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively. [...] hmm...
\begin{align}
	p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi) &= p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi) \times \det{\bf J}^{(i)}\\
	&= \left(\mathcal{N}(\W^{(i)} \given \W_s, \V_s) \times \det{\bf J}^{(i)}\right)\rvert_{t=\tau^{(i)}}
\end{align}

>>>>

Take ${\bf D}$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take ${\bf W}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame. Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy. The position of the satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, where the subscript $s$ refers to the satellite. The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$. Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).

We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$. In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref. response? refer to another section]. With these assumptions we can write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
\begin{align}
	%p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
	%	\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi)
	p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
		\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi)
\end{align}
where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters. We assume that the observational errors are Gaussian:
\begin{align}
	p(\D^{(i)} \given \W^{(i)}) &= \mathcal{N}(\W^{(i)} \given \D^{(i)}, \bSigma^{(i)})\\
	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bSigma_s)
\end{align}
where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively. [...] hmm...
\begin{align}
	p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi) &= p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi) \times \det{\bf J}^{(i)}\\
	&= \left(\mathcal{N}(\W^{(i)} \given \W_s, \V_s) \times \det{\bf J}^{(i)}\right)\rvert_{t=\tau^{(i)}}
\end{align}
